---
title: "HW1 SVM and KNN"
author: "Richard"
date: "8/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE}
library(kernlab)
library(ggplot2)
library(kknn)
setwd("C:/Users/richa/Documents/GT_Fall_2019/ISYE_6501/HW1")
data_df = read.table("credit_card_data-headers.txt", header = TRUE)
data = as.matrix(data_df)

model_func = function(C, kern, scaling){
  model <- ksvm(data[,1:10],data[,11],type="C-svc", kernel=kern,C=C,scaled=scaling)
  
  a = colSums(model@xmatrix[[1]] * model@coef[[1]])
  #a
  a0 = -model@b
  
  pred = predict(model,data[,1:10])
  acc = sum(pred == data[,11]) / nrow(data) * 100
  return(acc)
}

#Generate the list of C values be order of magnitude
clist = c()
for (i in -2:6){
  clist = c(clist, 10^i)
}

acc_v_T = c()
acc_v_F = c()
acc_p_T = c()
acc_p_F = c()

model1 = 1
if (model1 == 1){
for (i in clist){
  val_v_T = model_func(i, "vanilladot", TRUE)
  val_v_F = model_func(i, "vanilladot", FALSE)
  val_p_T = model_func(i, "polydot", TRUE)
  val_p_F = model_func(i, "polydot", FALSE)
  acc_v_T = c(acc_v_T, val_v_T)
  acc_v_F = c(acc_v_F, val_v_F)
  acc_p_T = c(acc_p_T, val_p_T)
  acc_p_F = c(acc_p_F, val_p_F)
}
}

# Create the accuracy df
acc_v_T_df = data.frame(c_vals=clist, accuracy=acc_v_T)
acc_v_F_df = data.frame(c_vals=clist, accuracy=acc_v_F)
acc_p_T_df = data.frame(c_vals=clist, accuracy=acc_p_T)
acc_p_F_df = data.frame(c_vals=clist, accuracy=acc_p_F)
```

We found that a good classfier for this dataset had the following equation:

It was observed that the coefficient for predictor A9 was much larger than the coefficients for all other predictor variables.
```{r}
model <- ksvm(data[,1:10],data[,11],type="C-svc", kernel="vanilladot",C=100,scaled=TRUE)
a = colSums(model@xmatrix[[1]] * model@coef[[1]])
a0 = -model@b
```

We used C=100, the vanilladot kernel, and scaled=TRUE after reviewing the effect of these parameters on model performance.  The scaled linear SVM model produced the highest accuracy in the C=1~250 range.
```{r}
  ggplot(data=acc_v_T_df, aes(x=c_vals, y=accuracy, group=1))+
    scale_x_log10()+
    geom_line()+
    geom_point()+
    labs(title="Accuracy of Linear SVM Over Range of C With Scaling",x="C Values", y = "Accuracy (%)")+
    theme_classic()
  
  ggplot(data=acc_v_F_df, aes(x=c_vals, y=accuracy, group=1))+
    scale_x_log10()+
    geom_line()+
    geom_point()+
    labs(title="Accuracy of Linear SVM Over Range of C Without Scaling",x="C Values", y = "Accuracy (%)")+
    theme_classic()
  
  ggplot(data=acc_p_T_df, aes(x=c_vals, y=accuracy, group=1))+
    scale_x_log10()+
    geom_line()+
    geom_point()+
    labs(title="Accuracy of Polynomial SVM Over Range of C With Scaling",x="C Values", y = "Accuracy (%)")+
    theme_classic()

  ggplot(data=acc_p_F_df, aes(x=c_vals, y=accuracy, group=1))+
      scale_x_log10()+
      geom_line()+
      geom_point()+
      labs(title="Accuracy of Polynomial SVM Over Range of C Without Scaling",x="C Values", y = "Accuracy (%)")+
      theme_classic()
```


