---
title: "HW_2.2"
author: "Arjun Goyal"
date: "8/28/2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kernlab)
library(kknn)
library(ggplot2)

cred_card_data_headers <- read.table('credit_card_data-headers.txt',header=TRUE)
```

## K-Nearest Neighbors 

#KKNN Function

We write the function *kknn_accuracy* to evaluate the k-nearest neighbors model using different values of k. 

```{r}

kknn_accuracy = function(k_val){
  pred_responses <- rep(0, nrow(cred_card_data_headers)) #vector of 0s to insert the predicted response value of the model.
  for (i in 1:nrow(cred_card_data_headers)){
    #using [-i] to exclude the ith data point from the nearest neighbor calculation
    kknn.model = kknn(R1 ~., train = cred_card_data_headers[-i,], test = cred_card_data_headers[i,], k = k_val, kernel = "rectangular",scale = TRUE)
    #fitted(kknn.model) gives the value of the fraction of nearest neighbors to the ith data point that are 1.
    fit <- fitted(kknn.model)
    binary_response <- as.integer(fit + 0.5) #maps the fitted value to 0 or 1
    pred_responses[i] <- binary_response #adds the model's fitted value to the pred_responses vector
  }
  acc = sum(pred_responses == cred_card_data_headers[,11]) / nrow(cred_card_data_headers)
  return(acc)
}

```

We run a for loop iterating through [0,20] for values of k.

```{r}
k_acc <- rep(0, 20)
for (i in 1:20){ #values of k
  k_acc[i] = kknn_accuracy(i)
}

```

To determine the best accuracy delivered by the model, we run:
```{r} 
max(k_acc) #best k accuracy
```

To determine the value of k for which this best accuracy occurs, we run the following code to find the index of the maximum accuracy:
```{r}
max_k <- which(k_acc %in% max(k_acc))
max_k
```

#Plotting Accuracies

We can create a plot of the values of k vs. the accuracies:
```{r, echo=FALSE}
ggplot(as.data.frame(k_acc), aes(x= 1:20, y=k_acc)) + geom_point() +
  ggtitle("Accuracies of K-Nearest Neighbors Model by Values of K") +
  xlab("Integer Values of K") + ylab("Accuracy")

```
