---
title: "HW5"
author: "Alan Lo, Arjun Goyal, James Trawick, Richard Hardis"
date: "9/20/2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ISYE 6501 HW5 Response

## Question 8.1

Describe a situation or problem from your job, everyday life, current events, etc., for which a linear regression model would be appropriate. List some (up to 5) predictors that you might use.

In daily life, one common activity for many people is working out.  A person who is trying to lose weight or track general health metrics may be interested in knowing how many calories he/she burns in a given period of time working out.  There are many factors that can be used to predict the total calorie expenditure of a workout for a given length.  Some example factors include resting metabolic rate, average heart rate over the workout period, a categorical predictor describing the type of exercises performed (bike vs swim vs bicep curls), temperature of the surrounding workout environment, bodyfat content, daily diet, and $VO_2 \:max$.

A linear regression model incorporating a combination of these factors and others could yield a useful prediction of caloric expenditure for a given workout.

## Using crime data from http://www.statsci.org/data/general/uscrime.txt  (file uscrime.txt, description at http://www.statsci.org/data/general/uscrime.html ), use regression (a useful R function is lm or glm) to predict the observed crime rate in a city with the following data:

M = 14.0
So = 0
Ed = 10.0
Po1 = 12.0
Po2 = 15.5
LF = 0.640
M.F = 94.0
Pop = 150
NW = 1.1
U1 = 0.120
U2 = 3.6
Wealth = 3200
Ineq = 20.1
Prob = 0.04
Time = 39.0

Show your model (factors used and their coefficients), the software output, and the quality of fit. 


```{r}
crime_df = read.table("uscrime.txt", header = TRUE)
model = lm(Crime ~ ., crime_df)
summary(model)
```

From the above R output from the lm linear modeling function, we can see all of the factors and their coefficients.  The adjusted R-squared value is 0.7078, which is a high value for human-related modeling scenarios.  Additionally, based on the data summary seen below, we can see that all of the values of the input data lie within the ranges of each factor.  

Although the dataset is small, potentially leading to overfitting, the high $R^2$ value along with the all of the input data not requiring extrapolation gives us confidence in the predictive capability of the model.

```{r}
summary(crime_df)
```


The predicted crime rate for the given factor values can be found as follows:

```{r}
new_data = data.frame(
M = 14.0,
So = 0,
Ed = 10.0,
Po1 = 12.0,
Po2 = 15.5,
LF = 0.640,
M.F = 94.0,
Pop = 150,
NW = 1.1,
U1 = 0.120,
U2 = 3.6,
Wealth = 3200,
Ineq = 20.1,
Prob = 0.04,
Time = 39.0
)

predict(model, new_data)
```

The crime rate with the given input values is 155 offenses per 100,000 population in 1960.
